\documentclass[12pt]{report}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{bm}
\usepackage{multirow,booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz} 
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{calc}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage[margin=3cm]{geometry}
\newlength{\tabcont}
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.05in}
\usepackage{empheq}
\usepackage{framed}
\usepackage[most]{tcolorbox}
\usepackage{xcolor}
\colorlet{shadecolor}{orange!15}
\parindent 0in
\parskip 12pt
\geometry{margin=1in, headsep=0.25in}
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{reg}{Rule}
\newtheorem{exer}{Exercise}
\newtheorem{note}{Note}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{braket}
\usepackage{hyperref}
\usepackage{algpseudocode}

\newcommand{\balpha}{{\bm \alpha}}
\newcommand{\bbeta}{{\bm \beta}}
\newcommand{\beeta}{{\bm \eta}}
\newcommand{\bgamma}{{\bm \gamma}}
\newcommand{\bGamma}{{\bm \Gamma}}
\newcommand{\bdelta}{{\bm \delta}}
\newcommand{\bxi}{{\bm \xi}}
\newcommand{\bXi}{{\bm \Xi}}
\newcommand{\bchi}{{\bm \chi}}
\newcommand{\btheta}{{\bm \theta}}
\newcommand{\bTheta}{{\bm \Theta}}
\newcommand{\blambda}{{\bm \lambda}}
\newcommand{\bLambda}{{\bm \Lambda}}
\newcommand{\bmu}{{\bm \mu}}
\newcommand{\bsigma}{{\bm \sigma}}
\newcommand{\bSigma}{{\bm \Sigma}}
\newcommand{\bphi}{{\bm \phi}}
\newcommand{\bPhi}{{\bm \Phi}}
\newcommand{\bpsi}{{\bm \psi}}
\newcommand{\bPsi}{{\bm \Psi}}
\newcommand{\bpi}{{\bm \pi}}
\newcommand{\bPi}{{\bm \Pi}}

\newcommand{\ba}{{\bm a}}
\newcommand{\bb}{{\bm b}}
\newcommand{\bc}{{\bm c}}
\newcommand{\bd}{{\bm d}}
\newcommand{\be}{{\bm e}}
\newcommand{\bg}{{\bm g}}
\newcommand{\bi}{{\bm i}}
\newcommand{\bj}{{\bm j}}
\newcommand{\bk}{{\bm k}}
\newcommand{\bn}{{\bm n}}
\newcommand{\bs}{{\bm s}}
\newcommand{\bt}{{\bm t}}
\newcommand{\bu}{{\bm u}}
\newcommand{\bv}{{\bm v}}
\newcommand{\bw}{{\bm w}}
\newcommand{\bx}{{\bm x}}
\newcommand{\by}{{\bm y}}
\newcommand{\bz}{{\bm z}}
\newcommand{\bA}{{\bm A}}
\newcommand{\bB}{{\bm B}}
\newcommand{\bC}{{\bm C}}
\newcommand{\bD}{{\bm D}}
\newcommand{\bE}{{\bm E}}
\newcommand{\bF}{{\bm F}}
\newcommand{\bG}{{\bm G}}
\newcommand{\bH}{{\bm H}}
\newcommand{\bN}{{\bm N}}
\newcommand{\bS}{{\bm S}}
\newcommand{\bT}{{\bm T}}
\newcommand{\bU}{{\bm U}}
\newcommand{\bV}{{\bm V}}
\newcommand{\bW}{{\bm W}}
\newcommand{\bX}{{\bm X}}
\newcommand{\bY}{{\bm Y}}
\newcommand{\bZ}{{\bm Z}}

\newcommand{\bZero}{{\bm 0}}
\newcommand{\bOne}{{\bm 1}}
\newcommand{\bTwo}{{\bm 2}}
\newcommand{\bThree}{{\bm 3}}
\newcommand{\bFour}{{\bm 4}}
\newcommand{\bFive}{{\bm 5}}
\newcommand{\bSix}{{\bm 6}}
\newcommand{\bSeven}{{\bm 7}}
\newcommand{\bEight}{{\bm 8}}
\newcommand{\bNine}{{\bm 9}}

\hypersetup{
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = blue, %Colour for external hyperlinks
linkcolor    = blue, %Colour of internal links
citecolor   = green %Colour of citations
}

\allowdisplaybreaks

\graphicspath{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Notes on statistics}
\date{\today}
\author{Justin L. Ripley \\ ripley@illinois.edu}

\maketitle

\tableofcontents

\abstract{These notes contain brief reviews of a fairly random assortment of topics in probability and (Bayesian) statistics.
The only thing that unifies these notes is that at some point I wanted to better understand something (either for work or out of interest), and I have found I often retain information more effectively if I write it down somewhere.
\textbf{These notes are not self contained, may contain factual errors, and remain a work in progress.} 
Please let me know if you find any errors, or if you find any part of these notes to be unclear!  

For a good overview of basic (mostly frequentist) statistics, I recommend \cite{wasserman2010statistics}. 
I try to cite sources whenever possible (whenever I can remember the source I learned something from), although the purpose of these notes
are to serve more as a statistics ``cheat sheet'' than a formal review.

The notation is: vectors/tensors are in \textbf{boldfont}. 
Indices are denoted with lower case latin letters, e.g. the $i^{th}$ component of the vector
$\bv$ is $\left(\bv\right)_i = v_i$. We typically do not use boldfont when we explicitely
write down indices.
Repeated indices are summed over.
Capital $P$ always represents a probability distribution, $\bx$ always
represents an instantiation of measured data, $\btheta$ always represents
model parameters. More generally, model parameters are represented by 
greek letters, while data is represented by latin letters. 
Random variables are always capatalized. Partial derivatives are denoted
by $\partial$, and covariant derivatives by $\nabla_i$ (for our purposes,
you can usually replace covariant derivatives with partial derivatives).

I thank Rohit Chandramouli for helpful conversations, and a lecture on model selection that inspired the creation of these notes. 
I thank Simone Mezzasoma for helpful comments that have led to a clearer presentation.

Copyright 2024 Justin Ripley. 
You may copy and distribute this document provided that you make no changes to it.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Overview of parametric Bayesian statistics}
\label{chap:bayesian-overview}
\input{Bayesian-Overview/Bayesian-Overview.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Model comparison}
\label{chap:model-comparison}
\input{Model-Comparison/Model-Comparison.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Forecasting}
\label{chap:forecasting}
\input{Forecasting/Forecasting.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Times series analysis}
\label{chap:time-series-analysis}
\input{Time-Series-Analysis/Time-Series-Analysis.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Markov Chains}
\label{chap:markov-chains}
\input{Markov-Chains/Markov-Chains.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Martingales}
\label{chap:martingales}
\input{Martingales/Martingales.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Numerical integration}
\label{chap:numerical-integration}
\input{Numerical-Integration/Numerical-Integration.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Numerical optimization}
\label{chap:numerical-optmization}
\input{Numerical-Optimization/Numerical-Optimization.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Linear regression}
\label{chap:linear-regression}
\input{Linear-Regression/Linear-Regression.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Probability theory}
\label{chap:probability-theory}
\input{Probability-Theory/Probability-Theory.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Common probability distributions}
\label{chap:probability-distributions}
\input{Probability-Distributions/Probability-Distributions.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Common special functions}
\label{chap:special-functions}
\input{Special-Functions/Special-Functions.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Fourier and other transforms}
\label{chap:fourier-transforms}
\input{Fourier-Transforms/Fourier-Transforms.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Stationary phase approximation}
\label{chap:stationary-phase}
\input{Stationary-Phase/Stationary-Phase.tex}

\cleardoublepage 
\bibliographystyle{alpha}  
\bibliography{thebib}  %assuming your bibtex file is thesis.bib

\end{document}
